---
title: "Pipe Dreams Are Made of These"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Dependency
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
```
### Read CSV file
```{r}
students <-read_csv("Resources/students.csv")
schools <- read_csv("Resources/schools.csv")
```
### Preview tibble
```{r}
students %>% head()
```
```{r}
schools %>% head()
schools
```

#### Join two tibbles
```{r}
data2 = left_join(students, schools, by=c("school_name"))
data2 %>% head()
```

### Total Number of Schools
```{r}
school_count <- students$school_name %>% 
  unique() %>% 
  length()

school_count
```

### Total Number of Students
```{r}
student_count <-  students %>% nrow()
student_count
```

### Average reading and math scores
```{r}
mean_reading_score <- summarize(students, mean(reading_score))
mean_math_score <- summarize(students, mean(math_score))
```
### Calculate the percentage of students with passing reading scores, i.e. over 70%.
```{r}
percentage_passing_reading <- students %>% 
  filter(reading_score > 70) %>% 
  nrow() * 100 / student_count %>% 
  round(2)

percentage_passing_reading
```
### Calculate the percentage of students with passing math scores, i.e. over 70%.
```{r}
percentage_passing_math <-  students %>% 
  filter(math_score > 70) %>% 
  nrow() * 100 / student_count %>% 
  round(2)

percentage_passing_math
```
### Calculate the overall passing rate, i.e. the average of math and reading passing percentages
```{r}
overall_passing_rate <- (percentage_passing_math + percentage_passing_reading) / 2
overall_passing_rate
```

### Calculate the average math and reading scores by school
```{r}
students %>% 
  group_by(school_name) %>% 
  summarize(avg.reading=mean(reading_score), avg.math=mean(math_score))
```

### Calculate the average math and reading scores by grade level at each school
```{r}
students %>% 
  group_by(school_name, grade) %>% 
  summarize(avg.reading=mean(reading_score), avg.math=mean(math_score))
```
```{r}
total_budget <- schools %>% 
  summarize(sum(budget))
```

### Create summary by school
```{r}
school.tb <- data2 %>% 
  group_by(type, school_name) %>% 
  summarise(Avg.Reading.Score=mean(reading_score), 
            Avg.Math.Score=mean(math_score), 
            Total.Students=n())
school.tb
```


### Display data
```{r}
paste("School count: ", school_count)
paste("Student count: ", student_count)
paste("Total budget: ", total_budget)
paste("Average reading score: ", mean_reading_score)
paste("Average math score: ", mean_math_score)
paste("% passing reading: ", percentage_passing_reading)
paste("% passing math: ", percentage_passing_math)
paste("Overall passing rate: ", overall_passing_rate)
```

### Use sapply() to convert data type
```{r}
budget_num <- total_budget %>% sapply(as.numeric)
reading_num <- mean_reading_score %>% sapply(as.numeric)
math_num <- mean_math_score %>% sapply(as.numeric)
```

### Create Tibble of District Summary
```{r}
district.tb <- tribble(
  ~Total.Schools, ~Total.Students, ~Total.Budget, ~Avg.Reading, ~Avg.Math, ~Percent.Pass.Reading, ~Percent.Pass.Math, ~Overall.Passing, 
  school_count, student_count, budget_num, reading_num, math_num, percentage_passing_reading, percentage_passing_math, overall_passing_rate)
```
### Display summary of district-wide data
```{r}
district.tb
```